---
title: "p8105_hw1_cl4775"
output: github_document
date: "2025-09-20"
---

```{r setup, include=FALSE}
library(tidyverse)
library(moderndive)
data("early_january_weather")
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

## Dataset Summary 

The dataset has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.  
The mean temperature is `r mean(early_january_weather$temp, na.rm = TRUE)`.

**Variables include:**

- `temp` – Temperature
- `humid` – Humidity
- `time_hour` – Hour of observation

```{r scatterplot, fig.width=6, fig.height=4}
ggplot(early_january_weather, aes(x = hour, y = temp, color = humid)) +
  geom_point() +
  labs(
  x = "Hour of Day",
  y = "Temperature",
  color = "Humidity"
  ) +
  theme_minimal()

ggsave("scatterplot_temp_vs_hour.png", width = 6, height = 4)
```


As we can see from the generated plot, in general, temperature rises in the middle of the day, and is lower at the beginning and end of days. 


# Problem 2

```{r}
df <- tibble(
  sample = rnorm(10),
  is_positive = sample > 0,
  chars = letters[1:10],
  factors = factor(rep(c("A","B","C"), length.out = 10))
)


#test mean 
mean(df %>% pull(sample))        # works
mean(df %>% pull(is_positive))   # works (TRUE -> 1, FALSE -> 0)
mean(df %>% pull(chars))         # returns NA with warning
mean(df %>% pull(factors))       # returns NA with warning

#test coercion
as.numeric(df$is_positive)  # TRUE=1, FALSE=0
as.numeric(df$chars)        # NA with warning
as.numeric(df$factors)      # integers representing factor levels

```

When trying to take the mean of data set, some variables work while others don't, since mean() only works on numeric or logical vectors. When the characters are not numbers, mean() cannot compute a value. Coersion helps demonstrate that variable types matter when performing calculations in R and for the wrong variable type an error would occur. 



